---
title: "EpiSODIC outbreak report"
format:
  html:
    embed-resources: true
    toc: false
    theme: cosmo
params:
  data_path: "report_data.rds"
---

<!--
# ===================================================================== #
#  An R package by Certe:                                               #
#  https://github.com/certe-medical-epidemiology                        #
#                                                                       #
#  Licensed as GPL-v2.0.                                                #
#                                                                       #
#  Developed at non-profit organisation Certe Medical Diagnostics &     #
#  Advice, department of Medical Epidemiology.                          #
#                                                                       #
#  This R package is free software; you can freely use and distribute   #
#  it for both personal and commercial purposes under the terms of the  #
#  GNU General Public License version 2.0 (GNU GPL-2), as published by  #
#  the Free Software Foundation.                                        #
#                                                                       #
#  We created this package for both routine data analysis and academic  #
#  research and it was publicly released in the hope that it will be    #
#  useful, but it comes WITHOUT ANY WARRANTY OR LIABILITY.              #
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(EpiSODIC)
library(ggplot2)
d <- readRDS(params$data_path)
obj <- d$obj
lang <- if (is.null(d$lang)) "nl" else d$lang
tr <- function(key, ...) episode_tr(key, lang = lang, ...)
na_dash <- function(x) if (is.na(x)) "-" else x
```

# `r episode_ui_italicise_taxon(obj$pathogen)` — `r tr(paste0("level.", obj$level))`

**`r obj$place`** &middot; `r tr("report.cluster")` #`r obj$id` &middot; `r obj$first_day` &ndash; `r obj$last_day`

`r tr("report.rendered_label")`: `r d$rendered_at` &middot; EpiSODIC `r d$package_version`

::: {.callout-note}
`r tr("report.note_versioned")`
:::

## `r tr("report.key_figures_title")`

| `r tr("report.col_metric")` | `r tr("report.col_value")` |
|---|---|
| `r tr("dossier.stat.observed")` | `r obj$n_cases` |
| `r tr("panel.trend.legend_expected")` | `r if (is.na(obj$expected)) tr("report.na") else round(obj$expected, 1)` |
| `r tr("dossier.stat.ratio")` | `r if (is.na(obj$ratio)) tr("report.na") else round(obj$ratio, 1)` |
| `r tr("dossier.stat.unique_patients")` | `r obj$unique_patients` (`r tr("report.isolates_suffix", n = obj$n_isolates)`) |
| `r tr("report.priority_score")` | `r round(obj$priority_score, 0)` |
| `r tr("report.detected_by")` | `r episode_ui_code_join(obj$detectors)` |

```{r curve-shape, results='asis'}
if (!is.null(obj$curve_shape) && !is.na(obj$curve_shape)) {
  cat("**", tr("report.curve_shape_label"), "** ", tr(paste0("interpretation.fragment.curve_shape.", obj$curve_shape)), "\n\n", sep = "")
}
```

## `r tr("panel.epicurve.title")`

```{r epi-curve, fig.width=7, fig.height=3}
if (!is.null(d$epi_curve) && nrow(d$epi_curve) > 0) {
  print(episode_ui_epi_curve_chart(d$epi_curve, lang = lang))
} else {
  cat(tr("report.no_data"))
}
```

```{r rt-section, results='asis'}
if (!is.null(obj$rt) && nrow(obj$rt) > 0) {
  cat("## ", tr("report.rt_title"), "\n\n", tr("report.rt_note"), "\n\n", sep = "")
}
```

```{r rt-plot, fig.width=7, fig.height=2.5}
if (!is.null(obj$rt) && nrow(obj$rt) > 0) {
  print(episode_ui_rt_chart(obj$rt, lang = lang))
}
```

```{r trend-section, results='asis'}
if (!is.null(d$trend) && nrow(d$trend) >= 4) {
  cat("## ", tr("panel.trend.title"), "\n\n", sep = "")
}
```

```{r trend-plot, fig.width=7, fig.height=2.5}
if (!is.null(d$trend) && nrow(d$trend) >= 4) {
  print(episode_ui_trend_chart(d$trend, lang = lang))
}
```

## `r tr("panel.geo.title")`

```{r geo, results='asis'}
if (!is.null(obj$concentration) && nrow(obj$concentration$rows) > 0) {
  rows <- obj$concentration$rows
  cat("|", tr("panel.geo.aside"), "|", tr("report.col_count"), "|\n|---|---|\n")
  for (i in seq_len(min(nrow(rows), 10))) {
    cat("|", rows$label[i], "|", rows$n[i], "|\n")
  }
  cat("\n*", tr("report.suppressed_note", threshold = d$small_count_threshold), "*\n", sep = "")
} else {
  cat(tr("panel.geo.empty"), "\n")
}
```

## `r tr("verloop.title")`

```{r timeline, results='asis'}
if (!is.null(d$timeline) && nrow(d$timeline) > 0) {
  for (i in seq_len(nrow(d$timeline))) {
    row <- d$timeline[i, ]
    cat("-", row$at, "&middot;", row$actor)
    if (!is.na(row$verdict_label)) cat(" &middot; **", row$verdict_label, "**", sep = "")
    if (row$kind == "closure") cat(" &middot;", tr("report.closed_note"))
    cat("\n")
    if (!is.na(row$rationale) && nzchar(row$rationale)) cat("  >", row$rationale, "\n")
  }
} else {
  cat(tr("report.timeline_empty"), "\n")
}
```

```{r similar-section, results='asis'}
if (!is.null(d$similar) && nrow(d$similar) > 0) {
  cat("## ", tr("panel.similar.title"), "\n\n", sep = "")
  cat("|", tr("panel.similar.col.place"), "|", tr("panel.similar.col.cases"), "|", tr("panel.similar.col.verdict"), "|\n|---|---|---|\n")
  for (i in seq_len(nrow(d$similar))) {
    row <- d$similar[i, ]
    cat("|", row$place, "|", row$n_cases, "|", ifelse(is.na(row$verdict_label), "-", row$verdict_label), "|\n")
  }
}
```

```{r linelist-section, results='asis'}
if (!is.null(d$linelist) && nrow(d$linelist) > 0) {
  cat("## ", tr("panel.linelist.title"), "\n\n", sep = "")
  cat("|", tr("panel.linelist.col.case"), "|", tr("panel.linelist.col.date"), "|", tr("panel.linelist.col.sex"), "|",
      tr("panel.linelist.col.age"), "|", tr("panel.linelist.col.pc4"), "|", tr("panel.linelist.col.ward"), "|",
      tr("panel.linelist.col.specialism"), "|\n")
  cat("|---|---|---|---|---|---|---|\n")
  for (i in seq_len(nrow(d$linelist))) {
    row <- d$linelist[i, ]
    cat("|", row$source_key, "|", row$sample_date, "|", na_dash(row$sex), "|", na_dash(row$age), "|",
        na_dash(row$pc4), "|", na_dash(row$ward), "|", na_dash(row$specialism), "|\n")
  }
}
```

---

*`r tr("report.footer")`*
