---
title: "Measuring detection against known truth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Measuring detection against known truth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

The dashboard's Performance screen reports how your board's own verdicts
fell out: of the clusters EpiSODIC put in front of them, how many were
judged real, and how long each took to assess. That is the right question
to ask of a running instance, and the wrong one to answer a different
question with. It is circular - it reports what the board thought of what
the system showed them - and sensitivity cannot be got out of it at all,
because an outbreak the detectors miss never becomes a cluster for
anybody to have an opinion about.

`episodic_validate_detection()` answers the other question. It measures
the detectors against outbreaks that are known to be there because they
were put there.

## What it does

```{r}
library(EpiSODIC)

result <- episodic_validate_detection()
result
```

Four things happen inside that call.

**It generates history.** `episodic_synthetic_cases()` produces years of
endemic laboratory data for a fictional region and injects six outbreaks
into it, each shaped for a different detection channel: a single case of
an always-notable pathogen, a three-case ward cluster, a nursing-home
outbreak, a tight point source, a propagated outbreak over four
generations, and a regional wave deliberately spread so thin that no
single place ever sees enough of it for a rule to notice.

**It replays that history one week at a time.** Each run gets only the
cases sampled on or before its own date - what a laboratory extract taken
that day would actually have contained. A single run over the whole
history would see an outbreak's last case at the same moment as its
first, which says nothing at all about timeliness.

**It matches on case sets, in both directions.** Not on overlapping
dates: an endemic winter cluster spans the same fortnight as a seeded
outbreak and shares not one case with it.

**It reports what it found**, with the denominators and the intervals
that belong to each number.

## The ground truth is data

```{r}
cases <- episodic_synthetic_cases(
  start_date = as.Date("2025-01-01"), end_date = as.Date("2025-12-31")
)
truth <- episodic_synthetic_ground_truth(cases)

truth$outbreaks[, c("outbreak_id", "pathogen", "n_cases", "first_day")]
head(truth$cases)
```

Every injected case also carries a `PT-OUTBREAK-*` patient key, and it is
tempting to recover membership by matching on it. Do not. That couples
whatever is doing the matching to a naming convention nobody has
undertaken to keep, and it fails quietly - as a smaller set of outbreak
cases, not as an error - the first time the convention changes.

## The two matching rules

An **outbreak is detected** when some cluster's final membership holds at
least `min_recall` of its cases. A **cluster is a true positive** when at
least `min_precision` of *its own* cases belong to one seeded outbreak.
A dossier that is one part outbreak to five parts background is a false
alarm, and counting it otherwise would be flattering rather than
measuring.

Both are choices, and a result that only holds at one particular pair of
values is not a result. `episodic_validate_rethreshold()` re-matches a
replay that has already been run, so sweeping them costs arithmetic
rather than another few hours of detection:

```{r}
strict <- episodic_validate_rethreshold(
  result,
  min_recall = 0.8, min_precision = 0.8
)
strict$summary[strict$summary$metric == "sensitivity", ]
```

## What it reports, and how to read it

`result$summary` is one row per (metric, group). Proportions -
sensitivity, PPV - carry a numerator, a denominator and a Wilson
interval. False alarms are a **rate** per stream-week, with the
stream-week denominator stated and an exact Poisson interval: a count of
false alarms without its denominator is uninterpretable, and here the
denominator is large and moves with the configuration.

```{r}
result$summary[result$summary$metric == "false_alarms", ]
result$meta$n_stream_weeks
```

Two conventions run through all of it.

**A metric with no denominator is `NA`, never 0.** A history with nothing
injected has no sensitivity; that is not a sensitivity of zero, and
averaging it as one would be a mistake.

**An outbreak nothing found is censored, not dropped.** A median delay
computed over the outbreaks that were detected is biased downward, and
the bias is worst exactly where the method is weakest, because the shapes
it misses are the slow ones. So the summary carries a Kaplan-Meier median
that keeps the misses in as right-censored observations, and the share
never detected beside the plain median. Quote them together, or quote the
censored one; never the plain median alone.

```{r}
prospective <- result$outbreaks[result$outbreaks$fully_prospective, ]
km <- episodic_validation_km(
  ifelse(prospective$detected, prospective$delay_from_first, prospective$censor_days),
  prospective$detected
)
```

## A comparator, and a control

Sensitivity and false-alarm figures mean little in isolation.
`episodic_validate_comparator()` runs a rule anyone could write in an
afternoon over exactly the same data, through exactly the same metrics:

```{r}
naive <- episodic_validate_comparator(method = "same_place")
shewhart <- episodic_validate_comparator(method = "shewhart")
```

And a negative control - the same endemic history with nothing seeded in
it - is the only clean specificity measurement available, because on
seeded data every "false" alarm might be a real cluster the generator
produced by chance:

```{r}
control <- episodic_validate_detection(outbreaks = FALSE)
control$summary[control$summary$metric == "alarms", ]
```

## Which detector earns its place

Which detector fired first answers "who got there first", not "what does
each channel add": two detectors may both find an outbreak, and the loser
looks worthless until you remove the winner. So the question is asked by
removing them, one at a time:

```{r}
without_mem <- episodic_validate_detection(
  detectors = c("farrington", "same_place", "rare_trigger")
)
```

`config$<detector>$enabled` is what that sets, and it is part of
`config_hash`, so two runs over the same data with different detectors
switched on can never come out looking identical.

## Where the real numbers live

The defaults above are deliberately small - one seed, a short window, a
single year of history - so that this returns in seconds and can be run
from an example. They are a demonstration of the machinery and not a
study: at one year of history the statistical detectors have no baseline
to fit against, and one seed is one realisation.

The study itself is `data-raw/validation/run_study.R`: twenty seeds by
twenty-six weekly runs, four years of history, a negative control, a
drop-one analysis for each of the four detectors, two comparators and an
operating-point sweep. It takes hours, it writes its results to
`data-raw/validation/results/` as CSV with the package version and
`config_hash` on every row, and it ships with nothing - `data-raw/` is in
`.Rbuildignore`, so none of it is in the built package, none of it runs
under `R CMD check`, and none of it is reachable from an installed
EpiSODIC.

If a number it produces is unflattering, that is a finding. Tuning the
shipped defaults until it looks better and then quoting it would be
fitting the evaluation to the answer.
