---
title: "Tuning detection for your catchment"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tuning detection for your catchment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This vignette is for someone with a running EpiSODIC instance whose detection behaviour does not quite match their catchment: too many false alarms, missing outbreaks, clusters closing too early, the wrong things at the top of the board. It maps each of those complaints to the parameters that control it, the trade-off each adjustment creates, and how to verify the change helped. It does not re-explain how the detectors work (`vignette("detection-reconciliation")`) or how the configuration overlay system works (`vignette("deployment")`).

## Before you change anything

Every detection run records the exact configuration it used: a SHA-1 hash of the resolved settings (`config_hash`) and a full JSON snapshot (`config_snapshot`), both on the run's own row. A change you make today affects only future runs, never the meaning of a past one. You can tune with confidence, since nothing becomes retroactively wrong.

Point the `EPISODIC_CONFIG` environment variable at a YAML file with only the keys you want to change; everything else keeps its shipped default. See `vignette("deployment")` for the mechanics.

```yaml
# my_instance.yaml - only what differs from the defaults.
farrington:
  alpha: 0.01
```

Three guidelines before you start:

1. Change one thing at a time. Two changes at once, and the effect
   could be either or both.
2. State what you expect the change to do before you make it.
3. Decide how you will measure whether it worked: the Performance
   screen for a live instance, or `episodic_validate_detection()` for a
   controlled experiment. See "Verifying that a change helped" below.


## Too many false alarms

### The Farrington bound is too narrow

If Farrington fires on modest rises that are within normal variation for your catchment, widen the statistical bound or raise the effect-size floor that gates new clusters.

`farrington.alpha` is the significance level for the upper bound. Lowering it from 0.05 to 0.01 widens the bound, so a run needs a larger exceedance before it reports one. This is a sensitivity-specificity trade-off: fewer false alarms, but genuine small outbreaks take longer to appear or are missed entirely.

`effect_size_floor.min_excess_over_upperbound` (default: 3) requires at least this many cases above the bound before a *new* cluster is opened. Raising it filters out borderline exceedances. A match against an already-open cluster is always merged, regardless of this floor.

`effect_size_floor.min_ratio_observed_expected` (default: 1.5) requires the observed/expected ratio to reach this level. Together with the excess floor, it means a signal must be both statistically and practically notable. Either can be set to `~` (null) to disable it.

```yaml
farrington:
  alpha: 0.01
effect_size_floor:
  min_excess_over_upperbound: 5
  min_ratio_observed_expected: 2.0
```

To measure the effect before deploying it:

```{r}
result_before <- episodic_validate_detection(history_years = 4)
result_after  <- episodic_validate_detection(
  history_years = 4,
  config = list(
    farrington = list(alpha = 0.01),
    effect_size_floor = list(min_excess_over_upperbound = 5)
  )
)
result_before$summary[result_before$summary$metric == "sensitivity", ]
result_after$summary[result_after$summary$metric == "sensitivity", ]
```

### same_place fires on endemic patterns

The rule-based detector opens clusters whenever N cases of the same pathogen appear at the same institution within K days. For a common organism at a large hospital, three cases in 14 days may be entirely unremarkable.

The global defaults (`same_place.default_n_cases` and `same_place.default_k_days`) affect every pathogen that does not have its own override. Raising `default_n_cases` globally also raises it for rare organisms where it should stay low, so per-pathogen overrides under `same_place.overrides` are usually the better lever:

```yaml
same_place:
  overrides:
    Escherichia coli: {n_cases: 5, k_days: 14}
    Staphylococcus aureus: {n_cases: 4, k_days: 10}
```

Lowering `default_k_days` requires cases to be more tightly clustered in time, which reduces noise from background endemic rates.

### Suppression lets redundant signals through

If the same real outbreak appears as separate clusters at multiple lattice levels (a ward cluster and its parent institution cluster), suppression is not aggressive enough.

`suppression.child_dominance_threshold` (default: 0.70) is the share of a parent's cases that a single child must account for to suppress the parent. Lowering it makes it easier for a child to absorb its parent. `suppression.parent_diffuse_threshold` (default: 0.50) and `suppression.parent_min_flagged_children` (default: 2) control the reverse: a parent absorbs its children when no single child dominates and enough children overlap. A cluster someone has already assessed is never suppressed, regardless of these thresholds.

```yaml
suppression:
  child_dominance_threshold: 0.60
  parent_min_flagged_children: 3
```

Too aggressive suppression hides a genuinely independent signal behind a coincidentally overlapping one. If you find suppressed clusters that should not have been, raise the thresholds.


## Missing outbreaks

### Farrington never fits

If Farrington reports nothing and the run log says a stream does not have enough baseline history, the model cannot fit because the stream's history is shorter than the baseline requirement.

`farrington.b` (default: 2) is the number of reference years. The model needs `(b + 1) * 52` weeks before it fits at all, so at `b: 2` a stream needs three years of history. Lowering it to 1 reduces the requirement to two years, at the cost of a less stable model. Raise it as history accrues: 5 is the conventional choice for a mature instance.

The three eligibility gates (`eligibility.min_baseline_weeks`, `eligibility.min_median_weekly_count`, `eligibility.min_nonzero_week_share`) determine whether a stream reaches the statistical detectors at all. Streams that fail still run through `same_place` and `rare_trigger`. Loosening these lets the model fit on sparser streams, where it may not be meaningful.

```yaml
farrington:
  b: 1
eligibility:
  min_baseline_weeks: 26
```

### same_place thresholds are too high

A ward-level cluster of 2 cases of a serious organism may go unreported because the global `default_n_cases` is 3. A per-pathogen override is the targeted fix:

```yaml
same_place:
  overrides:
    Legionella: {n_cases: 2, k_days: 14}
```

### The eligibility gate excludes a stream

A stream you expect to be watched may fail the eligibility check. `eligibility.min_nonzero_week_share` (default: 0.2) requires at least 20% of baseline weeks to have at least one case. `eligibility.min_median_weekly_count` (default: 1) requires a median weekly count of at least 1. For genuinely sparse organisms, lowering these lets the stream through, though the statistical model is not meaningful on a stream with very few cases.

```yaml
eligibility:
  min_nonzero_week_share: 0.10
  min_median_weekly_count: 0
```


## Clusters close too early or too late

### Clusters close before the outbreak is really over

If a cluster auto-closes and a new one opens for the same ongoing outbreak, the case-free interval or the run-count limit is too short.

`reconciliation.case_free_days_default` (default: 14) is the fallback interval: if no case arrives within this many days, the cluster's episode ends. Per-pathogen `case_free_days` in the pathogen configuration CSV overrides this for pathogens with longer incubation periods. `reconciliation.close_after_runs` (default: 14) closes an unconfirmed cluster after this many consecutive runs with no new case.

```yaml
reconciliation:
  case_free_days_default: 21
  close_after_runs: 21
```

A cluster that has been confirmed by an epidemiologist is never auto-closed by `close_after_runs`, only by case-free days.

### Clusters stay open long after the last case

If the active queue has stale clusters nobody has looked at, `reconciliation.stale_open_days` (default: 60) force-closes any still-unassessed cluster this many days after its last case. `reconciliation.autoclose_unassessed` must be `true` (the default) for either `close_after_runs` or `stale_open_days` to act on an unassessed cluster.

```yaml
reconciliation:
  stale_open_days: 30
  close_after_runs: 7
```

Aggressive auto-closure can close a genuine slow-burn outbreak before anyone has assessed it.

### Cooldown absorption swallows a real recurrence

After a cluster closes, a cooldown window (per-pathogen `cooldown_days` in the pathogen configuration CSV) absorbs nearby cases into the recently-closed cluster rather than opening a new one. `reconciliation.cooldown_reopen_ratio` (default: 1.5) is an escape hatch: if the absorbed cases are at least 1.5 times the cluster's own size, the absorption is flagged rather than silent. Set to `~` to disable this escape hatch entirely, or shorten `cooldown_days` per pathogen to narrow the absorption window.

```yaml
reconciliation:
  cooldown_reopen_ratio: ~
```


## Priority ranking does not match your triage

The priority score is a 0-100 weighted mean of seven components. A component that cannot be computed for a given cluster (for example, `density_component` requires a patient-day denominator) drops out and the remaining weights renormalise.

`priority_score.weights` controls how much each component contributes:

```yaml
priority_score:
  weights:
    excess_component: 0.25
    ratio_component: 0.15
    severity_component: 0.20
    growth_component: 0.15
    agreement_component: 0.10
    density_component: 0.10
    spatial_component: 0.05
```

Common adjustments:

- **Low-severity organisms rank too high**: raise `severity_component`.
  It reads from `severity_weight` in the pathogen configuration CSV.
- **Large-count clusters of common organisms outrank small-count
  clusters of serious ones**: raise `severity_component`, lower
  `excess_component`.
- **Spatially concentrated clusters should rank higher**: raise
  `spatial_component`.
- **Clusters where only one detector fired rank too high**: raise
  `agreement_component`, so signals corroborated by multiple detectors
  are ranked above those seen by one alone.

All seven weights must sum to 1.0.


## Seasonal detection is wrong

### MEM never runs

If no epidemics appear even for a pathogen you expect to be seasonal, MEM may lack enough history to fit.

`mem.min_seasons` (default: 2) requires two fully-observed prior seasons before fitting. `mem.min_climatology_years` (default: 3) requires three complete ISO years of case history to derive a season anchor. Both are "need real history first" dials with the same consequence as `farrington.b`: set beyond what an instance carries, MEM says nothing. Lower them for a newer instance, and raise them as history accrues.

Per-pathogen `mem_mode` in the pathogen configuration CSV overrides the automatic seasonality test. Set to `yes` to force MEM for a pathogen that the automatic test (`auto`) does not flag as seasonal, or to `no` to suppress it for one that it does.

```yaml
mem:
  min_seasons: 1
  min_climatology_years: 2
```

### MEM fires too early or too late

If the pre-epidemic threshold crossing does not align with clinical reality, adjust the climatology or the seasonality test.

`mem.climatology_smooth_weeks` (default: 5) is the smoothing window for the 52-week climatology used to find the season trough. `mem.trough_percentile` (default: 0.25) is the quantile at or below which a week counts as part of the trough. `mem.seasonality_peak_weeks` (default: 8) and `mem.seasonality_min_peak_share` (default: 0.40) control the automatic seasonality test: a pathogen is seasonal under `mem_mode = 'auto'` if at least 40% of its detrended annual cases fall within an 8-week peak window.

```yaml
mem:
  climatology_smooth_weeks: 7
  trough_percentile: 0.20
```

Declaring the season started or ended is always a human act, never automatic. MEM raises the signal; a person declares. See `vignette("detection-reconciliation")`.


## Per-pathogen parameters

Per-pathogen biology lives in a CSV, separate from the YAML configuration. The default file is `r EpiSODIC:::doc_system_file("inst/config/episodic_default_pathogen_config.csv")`.

To override it, point `EPISODIC_PATHOGEN_CONFIG` at your own CSV containing only the pathogens and columns you want to change. For each pathogen you list, non-empty values override the default; empty cells keep it. Pathogens not in your file keep their row entirely. New pathogens are added.

```r
Sys.setenv(EPISODIC_PATHOGEN_CONFIG = "/path/to/my_pathogens.csv")
```

A minimal overlay that changes only the severity weight for MRSA and adds a local pathogen:

```
pathogen,severity_weight,case_free_days
MRSA,0.95,
My Local Pathogen,0.60,21
```

The columns and what they control:

| Column                             | Effect                                                                                                         |
| ---------------------------------- | -------------------------------------------------------------------------------------------------------------- |
| `pathogen`                         | Match key, raw lab-provided string                                                                             |
| `episode_days`                     | Deduplication window for repeated isolates from the same patient (via `AMR::get_episode()`)                    |
| `incub_min_days`, `incub_max_days` | Incubation period bounds, shown on the dossier timeline                                                        |
| `case_free_days`                   | Days without a new case before a cluster's episode boundary, overrides <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;case_free_days_default</code> |
| `cooldown_days`                    | Cool-down window after closure during which new cases are absorbed rather than opening a new cluster           |
| `rt_applicable`                    | Whether Rt estimation applies (0 or 1)                                                                         |
| `si_mean_days`, `si_sd_days`       | Serial interval parameters for Rt estimation                                                                   |
| `si_dist`                          | Serial interval distribution: `gamma`, `lognormal`, or `weibull`                                               |
| `mem_mode`                         | MEM seasonal detection: `auto` (derive from data), `yes` (force on), `no` (force off)                          |
| `severity_weight`                  | Clinical severity weight (0 to 1) feeding the priority score's severity component                              |

A common confusion point: the per-pathogen *detection* thresholds for `same_place` are in the YAML (under `same_place.overrides`), not in this CSV. The CSV holds per-pathogen *biology*, the YAML holds per-pathogen *detection rules*.

The validation framework accepts per-pathogen overrides through the `pathogen_config` argument:

```{r}
result <- episodic_validate_detection(
  history_years = 4,
  pathogen_config = data.frame(
    pathogen = "MRSA",
    severity_weight = 0.95,
    stringsAsFactors = FALSE
  )
)
```


## Verifying that a change helped

### Against synthetic ground truth

`episodic_validate_detection()` replays generated history through the full detection pipeline week by week and matches what it found against seeded outbreaks on case sets. Pass a `config` list (for YAML keys) or a `pathogen_config` data frame (for CSV columns) with the settings you want to test:

```{r}
baseline <- episodic_validate_detection(
  seeds = 1:10,
  history_years = 4,
  evaluation_weeks = 52
)

adjusted <- episodic_validate_detection(
  seeds = 1:10,
  history_years = 4,
  evaluation_weeks = 52,
  config = list(farrington = list(alpha = 0.01))
)

# Compare
baseline$summary[baseline$summary$metric %in% c("sensitivity", "ppv"), ]
adjusted$summary[adjusted$summary$metric %in% c("sensitivity", "ppv"), ]
```

This measures the machinery against known truth: does the change improve sensitivity, reduce false alarms, or affect time-to-detection? The defaults are deliberately small (one seed, four weeks) so it returns in seconds; for a proper comparison use more seeds and a longer window. See `vignette("detection-validation")` for the full framework.

### Against your own board's verdicts

The Performance screen measures the instance against its own epidemiologists' verdicts: what share of clusters the board classified as noise (positive predictive value), and how quickly real signals were detected (time-to-detection). This is not circular for the question "did this change reduce false alarms", which is exactly what PPV answers. It cannot measure sensitivity, because an outbreak the detectors miss never becomes a cluster for anyone to judge.

`config_hash` lets you compare runs before and after a change: filter the Performance screen or the `episodic_detection_run` table by hash to see the metrics under the old settings beside the new ones.


## Reference

Every tunable parameter across both the YAML configuration and the per-pathogen CSV, with its default and one-line effect.

### Reconciliation

| Key                                     | Default | Effect                                                                        |
| --------------------------------------- | ------: | ----------------------------------------------------------------------------- |
| <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;case_free_days_default</code> | 14      | Fallback days without a case before a cluster episode boundary                |
| <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;close_after_runs</code>       | 14      | Consecutive runs with no new case before an unconfirmed cluster auto-closes   |
| <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;cooldown_reopen_ratio</code>  | 1.5     | Absorption escape-hatch ratio; `~` disables                                   |
| <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;stale_open_days</code>        | 60      | Force-closes unassessed clusters this many days after last case; `~` disables |
| <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;autoclose_unassessed</code>   | `TRUE`  | Whether auto-closure rules apply to clusters nobody assessed                  |

### Eligibility

| Key                                   | Default | Effect                                                                            |
| ------------------------------------- | ------: | --------------------------------------------------------------------------------- |
| <code style="white-space: nowrap;">eligibility:<br>&nbsp;&nbsp;min_baseline_weeks</code>      | 52      | Weeks of history required before statistical detection runs                       |
| <code style="white-space: nowrap;">eligibility:<br>&nbsp;&nbsp;min_median_weekly_count</code> | 1       | Median weekly count over baseline; streams below are excluded from Farrington/MEM |
| <code style="white-space: nowrap;">eligibility:<br>&nbsp;&nbsp;min_nonzero_week_share</code>  | 0.2     | Share of baseline weeks with at least one case                                    |

### Effect-size floor

| Key                                             | Default | Effect                                                                            |
| ----------------------------------------------- | ------: | --------------------------------------------------------------------------------- |
| <code style="white-space: nowrap;">effect_size_floor:<br>&nbsp;&nbsp;min_excess_over_upperbound</code>  | 3       | Cases above Farrington's upper bound required to open a new cluster; `~` disables |
| <code style="white-space: nowrap;">effect_size_floor:<br>&nbsp;&nbsp;min_ratio_observed_expected</code> | 1.5     | Observed/expected ratio required to open a new cluster; `~` disables              |

### same_place detector

| Key                               | Default        | Effect                                                          |
| --------------------------------- | -------------: | --------------------------------------------------------------- |
| <code style="white-space: nowrap;">same_place:<br>&nbsp;&nbsp;enabled</code>              | `TRUE`         | Master switch                                                   |
| <code style="white-space: nowrap;">same_place:<br>&nbsp;&nbsp;default_n_cases</code>      | 3              | Cases at one institution within K days to fire                  |
| <code style="white-space: nowrap;">same_place:<br>&nbsp;&nbsp;default_k_days</code>       | 14             | Time window in days                                             |
| <code style="white-space: nowrap;">same_place:<br>&nbsp;&nbsp;lookback_days</code>        | 90             | How far back a run reports hits; `~` unbounds (not recommended) |
| <code style="white-space: nowrap;">same_place:<br>&nbsp;&nbsp;overrides:<br>&nbsp;&nbsp;&nbsp;&nbsp;&lt;pathogen&gt;</code> | (per pathogen) | `{n_cases, k_days}` per raw pathogen string                     |

### Farrington detector

| Key                               | Default | Effect                                                                              |
| --------------------------------- | ------: | ----------------------------------------------------------------------------------- |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;enabled</code>              | `TRUE`  | Master switch; switching off stops the detector, not the Pathogen screen's baseline |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;aggregation</code>          | week    | Aggregation unit                                                                    |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;b</code>                    | 2       | Reference years; model needs `(b+1)*52` weeks                                       |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;w</code>                    | 3       | Half-window (weeks) around the same calendar week in each reference year            |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;reweight</code>             | `TRUE`  | Down-weight past outbreak weeks by Anscombe residual                                |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;weightsThreshold</code>     | 2.58    | Residual cutoff for reweighting                                                     |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;trend</code>                | `TRUE`  | Allow a linear trend term                                                           |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;pastWeeksNotIncluded</code> | 26      | Trailing weeks excluded from the noise estimate                                     |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;limit54</code>              | [5, 4]  | Low-count safeguard                                                                 |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;alpha</code>                | 0.05    | Significance level; lower = wider bound = fewer alarms                              |
| <code style="white-space: nowrap;">farrington:<br>&nbsp;&nbsp;max_weeks_tested</code>     | 8       | Cap on trailing weeks tested per run                                                |

### MEM detector

| Key                              | Default                              | Effect                                                                      |
| -------------------------------- | -----------------------------------: | --------------------------------------------------------------------------- |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;enabled</code>                    | `TRUE`                               | Master switch                                                               |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;min_seasons</code>                | 2                                    | Fully-observed prior seasons required before fitting                        |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;levels</code>                     | [pathogen_province, pathogen_region] | Lattice levels MEM runs at                                                  |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;min_climatology_years</code>      | 3                                    | Complete ISO years required to derive a season anchor                       |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;climatology_smooth_weeks</code>   | 5                                    | Smoothing window for the 52-week climatology                                |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;trough_percentile</code>          | 0.25                                 | Quantile at or below which a week is part of the trough                     |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;seasonality_peak_weeks</code>     | 8                                    | Peak-concentration window width for the seasonality test                    |
| <code style="white-space: nowrap;">mem:<br>&nbsp;&nbsp;seasonality_min_peak_share</code> | 0.40                                 | Minimum share of detrended cases in the peak window for `mem_mode = 'auto'` |

### rare_trigger detector

| Key                          | Default     | Effect                                                          |
| ---------------------------- | ----------: | --------------------------------------------------------------- |
| <code style="white-space: nowrap;">rare_trigger:<br>&nbsp;&nbsp;enabled</code>       | `TRUE`      | Master switch                                                   |
| <code style="white-space: nowrap;">rare_trigger:<br>&nbsp;&nbsp;pathogens</code>     | (4 shipped) | Curated list of raw pathogen strings; a single case fires       |
| <code style="white-space: nowrap;">rare_trigger:<br>&nbsp;&nbsp;min_cases</code>     | 1           | Cases in the window to fire                                     |
| <code style="white-space: nowrap;">rare_trigger:<br>&nbsp;&nbsp;lookback_days</code> | 90          | How far back a run reports hits; `~` unbounds (not recommended) |

### Priority score

| Key                                          | Default | Effect                                                      |
| -------------------------------------------- | ------: | ----------------------------------------------------------- |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;excess_component</code>    | 0.25    | Weight for cases above the model's upper bound              |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;ratio_component</code>     | 0.15    | Weight for observed/expected ratio                          |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;severity_component</code>  | 0.20    | Weight for the pathogen's severity_weight                   |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;growth_component</code>    | 0.15    | Weight for cluster acceleration                             |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;agreement_component</code> | 0.10    | Weight for share of detectors that fired                    |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;density_component</code>   | 0.10    | Weight for incidence per 1,000 patient-days versus baseline |
| <code style="white-space: nowrap;">priority_score:<br>&nbsp;&nbsp;weights:<br>&nbsp;&nbsp;&nbsp;&nbsp;spatial_component</code>   | 0.05    | Weight for case concentration in one location               |

### Scale

| Key                     | Default                              | Effect                                                      |
| ----------------------- | ------------------------------------ | ----------------------------------------------------------- |
| <code style="white-space: nowrap;">scale:<br>&nbsp;&nbsp;epidemic_levels</code> | [pathogen_province, pathogen_region] | Lattice levels that produce epidemics rather than outbreaks |

### Geography

| Key                            | Default | Effect                                                |
| ------------------------------ | :------ | ----------------------------------------------------- |
| <code style="white-space: nowrap;">geography:<br>&nbsp;&nbsp;region_code</code>        | REGION  | L5 whole-catchment code; name your service area       |
| <code style="white-space: nowrap;">geography:<br>&nbsp;&nbsp;area_code_prefix</code>   | AREA-   | L3 prefix before the leading postcode characters      |
| <code style="white-space: nowrap;">geography:<br>&nbsp;&nbsp;area_pc_characters</code> | 2       | Number of leading postcode characters for L3 grouping |

### Suppression

| Key                                       | Default | Effect                                                     |
| ----------------------------------------- | ------: | ---------------------------------------------------------- |
| <code style="white-space: nowrap;">suppression:<br>&nbsp;&nbsp;child_dominance_threshold</code>   | 0.70    | Case-overlap share for a child to suppress its parent      |
| <code style="white-space: nowrap;">suppression:<br>&nbsp;&nbsp;parent_diffuse_threshold</code>    | 0.50    | Below this per child, the rise counts as diffuse           |
| <code style="white-space: nowrap;">suppression:<br>&nbsp;&nbsp;parent_min_flagged_children</code> | 2       | Minimum overlapping children before the diffuse rule fires |

### Report

| Key                            | Default | Effect                                                                                        |
| ------------------------------ | ------: | --------------------------------------------------------------------------------------------- |
| <code style="white-space: nowrap;">report:<br>&nbsp;&nbsp;small_count_threshold</code> | 5       | Counts below this in geography/institution tables are suppressed to prevent re-identification |
| <code style="white-space: nowrap;">report:<br>&nbsp;&nbsp;output_dir</code>            | ~       | Base directory for rendered reports; mandatory when `EPISODIC_DB` is a DSN                    |

### Access

| Key                    | Default | Effect                                                    |
| ---------------------- | ------- | --------------------------------------------------------- |
| <code style="white-space: nowrap;">access:<br>&nbsp;&nbsp;require_login</code> | `FALSE` | Whether the dashboard requires a sign-in to read anything |

### Per-pathogen CSV columns

| Column                             | Typical default | Effect                                            |
| ---------------------------------- | --------------: | ------------------------------------------------- |
| `episode_days`                     | 30              | Deduplication window for repeated isolates        |
| `incub_min_days`, `incub_max_days` | (varies)        | Incubation period bounds for the dossier timeline |
| `case_free_days`                   | 14              | Overrides <code style="white-space: nowrap;">reconciliation:<br>&nbsp;&nbsp;case_free_days_default</code> |
| `cooldown_days`                    | (varies)        | Cool-down absorption window after closure         |
| `rt_applicable`                    | 0               | Whether Rt estimation applies (0 or 1)            |
| `si_mean_days`, `si_sd_days`       | (varies)        | Serial interval for Rt                            |
| `si_dist`                          | (varies)        | Serial interval distribution                      |
| `mem_mode`                         | auto            | MEM: `auto`, `yes`, or `no`                       |
| `severity_weight`                  | 1.00            | Clinical severity for priority scoring (0 to 1)   |


## See also

- [**Detection and reconciliation**](https://certe-medical-epidemiology.github.io/EpiSODIC/articles/detection-reconciliation.html) - how the detectors, reconciliation, suppression, and MEM work
- [**Deployment**](https://certe-medical-epidemiology.github.io/EpiSODIC/articles/deployment.html) - the configuration overlay system and instance setup
- [**Measuring detection against known truth**](https://certe-medical-epidemiology.github.io/EpiSODIC/articles/detection-validation.html) - measuring detection against known truth
- [**Environment variables**](https://certe-medical-epidemiology.github.io/EpiSODIC/articles/environment-variables.html) - all `EPISODIC_*` variables including `EPISODIC_PATHOGEN_CONFIG`
